---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis 


<style>
p.blue_box {
  background-color: #4955a1;
  color: #FFFFFF;
  padding: 10px;
  border: 1px solid blue;
  margin-left: 25px;
  border-radius: 5px;
  font-style: normal;
}

p.heading {
  background-color: #a19549;
  margin: auto;
  width: 80%;
  padding: 5px;
  font-style: bold;
}

p.regular {
  margin: auto;
  padding: 5px;
}

</style>

<p class="blue_box">
Data Science projects begin with the problem at hand, and then we examine the available data. The available data should now assist us in addressing the problem. If the data itself is incorrect or insufficient, we must take a pause and either begin collecting additional data, new and relevant data, or an entirely new dataset in order to solve the problem. To determine if the data at hand is accurate, we will begin with descriptive statistics to determine the range of data values, mean, median, etc., followed by exploratory data analysis to determine the relationship between variables. Various techniques are employed during this analysis, beginning with the creation of visualizations, data cleansing to remove missing values, imputation of missing data with meaningful and close estimates, etc. </p>

<br>

<p class="heading"> <b>Problem statement: Who are the prospective customers from the available dataset who do not have health insurance and who should be contacted by our marketing team?</b> </p>

</br>

<p class="regular"> For this analysis, the data is assumed to be stored in a single data frame. But, generally, the data is spread across multiple tables in a normalized fashion (We are talking about the database normalization).  

The original dataset used for this exploratory analysis is available at the "Practical Data Science with R" author's primary [github repository](https://github.com/WinVector/PDSwR2)
</p>

```{r read_data, echo=F}
file_location <- paste0(getwd(), "/../Input/custdata.rds")
customer_data <- readRDS(file_location)
```

- Pay special attention to the missing values in the summary

```{r summarize_data}
summary(customer_data)
```

```{r handling_missing_data}
temp <- c(1, NA, -3, 5)
is.na(temp)
```

- Now looking at the columns (from customer_data) that have huge variation, can we explore these columns for outliers? 
- For example, the mean value of gas_usage is 41 and the 3rd quartile has 60. However, there is a max value of 570. Do you think there are any outliers in the data? 
- Plot a boxplot and see the outliers? 

```{r boxplot1}
library(ggplot2)
ggplot(data = customer_data) +
  geom_boxplot(mapping = aes(x = gas_usage))
```

- What is the spread of income? Is it normally spread? Draw a histogram. 
- set a limit on the income and see if the histogram looks better 
- You can include an income range from 0 to $200,000 
- Use scales::dollar_format() to include the labels in dollars 

```{r histogram}
layer1<- ggplot(data = customer_data)
layer1 +
  geom_histogram(aes(x=income)) +
  lims(x = c(0, 200000)) +
  scale_x_continuous(labels = scales::dollar_format(),
                     limits = c(0, 200000))
```

Conclusion: The data is not normally distributed.

- Density plots are another method for analyzing the distribution of univariate data  
- Use the density plot to examine the age range of the customers.  

```{r density_plot}
layer1 +
  geom_density(aes(x=age))

``` 

Analysis: the maximum number of customers in our data set are within the age of 25 to 75.

... Might help sometimes

```{r density_plot_log10}
layer1 +
  geom_density(aes(x=age)) +
  scale_x_log10()

```



- Create a bar chart to present the count values of various marital_statuses  
- Compare a stacked bar_plot against a side-by-side bar plot
- It would have been nice if there are proportion values to present the bar plots. This can be easily done with the dplyr and tidyr functions  

```{r bar_chart}
layer1 +
  geom_bar(aes(x = marital_status, fill = health_ins), stat = "count") +
  geom_text(aes(x = marital_status, label = ..count..),
            stat = "count")
  
```

```{r side-by-side_bar_plot}
ggplot(data = customer_data) + 
  geom_bar(aes(x = marital_status, fill = health_ins), stat = "count", position = "dodge") +
  geom_text(aes(x = marital_status, y = ..count.., label = ..count..),
            stat = "count", position = position_stack(vjust = 0.5), size = 3) +
  labs(title = "Marital Status and Health Insurance", x = "Marital Status", y = "Count", fill = "Health Insurance")

```


- We need count of each combination 

```{r aggregate_counts}
health_ins_count_by_marital_status <- aggregate(sex~marital_status+as.factor(health_ins), 
                                                data = customer_data, length)
names(health_ins_count_by_marital_status) <- c("marital_status","health_ins", "ct")
```

- Use the "ct" variable generated above to show the values as labels for the bar graph

```{r bar_graph_counts}
ggplot(health_ins_count_by_marital_status, aes(x=marital_status, y = ct)) +
  geom_bar(aes(fill=health_ins), stat = "identity") +
  geom_text(aes(label = ct), vjust = -1)
```

```{r side-by-side_bar_plot_2}
ggplot(health_ins_count_by_marital_status, aes(x=marital_status, y = ct)) +
  geom_bar(aes(fill=health_ins), stat = "identity", position = "dodge") +
  geom_text(aes(label = ct), position = position_dodge(width = 0.9), vjust = -1)
```


- Create a bar chart to present the count values of various state_of_res 
```{r bar_chart_desc}
layer1 +
  geom_bar(aes(x = reorder(state_of_res, state_of_res, FUN = function(x) length(x)))) +
  scale_x_discrete(name = "state of residence") +
  coord_flip()
```

- Compare marital status against housing type: Categorical vs. Categorical comparison 
- Facet plots with bar charts 

```{r facet_bar_plots}
ggplot(customer_data[!is.na(customer_data$housing_type),]) +
  geom_bar(aes(x = reorder(marital_status, marital_status, function(x) {length(x)}))) + 
  facet_wrap(~housing_type, scale = "free_x") +
  scale_x_discrete(name = "Marital Status") +
  coord_flip()

```

```{r facet_bar_plot2}
ggplot(data = customer_data[!is.na(customer_data$housing_type),]) +
  geom_bar(aes(x= marital_status, fill = health_ins), position = "dodge", stat = "count") +
  facet_wrap(~housing_type, scales = "free_x") +
  coord_flip()
```

- Compare population densities across categories 

```{r density_comparison_2}
ggplot(data = customer_data) +
  geom_density(aes(x = age, color = marital_status, linetype = marital_status, position="fill")) + 
  scale_color_brewer(palette="Dark2")
```

- Compare density plots for income and marital status

```{r density_comparison}
layer1 +
  geom_density(aes(x = income, color = marital_status, linetype = marital_status)) +
  scale_x_continuous(limits = c(0, 100000))
```

- Scatter plots to identify relationships between two continuos variables. 
- Create a scatter plot between age and income to see if there is any relationship between them or they are just independent variables 

```{r scatter-plot1}
ggplot(customer_data) + 
  geom_point(aes(x = age, y = income)) + 
  ggtitle(label = "Correlation analysis between Age and Income") + 
  geom_smooth(aes(x = age, y = income), method = "auto") + 
  lims(y = c(0, 200000))
```

```{r scatter-plot2}
layer1 +
  geom_point(aes(x = age, y = income)) +
  geom_smooth(aes(x = age, y = income, method = "auto"))
```

Conclusion: There is an increase in income among age groups between 25 and 60, and after that the income is flat.

<p class="blue_box">
In summary, the two dimensional visualizations provide a great starting point for exploratory data analysis. The below table summarizes  </p>


- There are so many customers with an income range of 0 - $100,000.   
- Plot a histogram to highlight customers with an income range 
  + -7000-20,000 (includes the min value)
  + 20,000 - 40,000
  + 40,000 - 60,000
  + 60,000 - 80,000 
  + 80,000 - 100,000
  + 100,000 - 200,000
  + 200,000 - 300,000 
  + 300,000 - 400,000
  + 400,000 - 1,300,000 (It includes a max value of 1,257,000)

- HINT: Use the "cut" function to get the custom breaks 

```{r histogram-custom-categories}
income_categories <- cut(customer_data$income, 
                          breaks = c(-7000, 20000, 40000, 60000, 80000, 100000, 200000, 300000, 400000, 1300000), 
                          labels = c("<20k", "20-40k", "40-60k", "60-80k", "80-100k", "100-200k", "200-300k", "300-400k", "400k-1.3M"),
                          na.rm = TRUE)

ggplot(data=customer_data, aes(x=income_categories)) +
  geom_histogram(color="white", fill="blue", stat = "count") +
  scale_x_discrete(name="Income", labels=c("<20k", "20-40k", "40-60k", "60-80k", "80-100k", "100-200k", "200-300k", "300-400k", "400k-1.3M"))

```


- More into modeling stuff later ...  
